Global search algorithm with local tuning for solving mixed-integer global optimization problems \thanks{This study was supported by the .}

### 2021-02-13 2021/MOTOR_BarkalovLebedev/MOTOR2021.tex
Motor исправление текста статьи

\begin{document}
%
\title{An Approach for Simultaneous Finding of Multiple Effective Decisions in Multi-objective Optimization Problems
%with non-convex constraints 
\thanks{This work was supported by the Ministry of Science and Higher Education of the Russian Federation, project no.0729-2020-0055, and by the Research and Education Mathematical Center, project no. 075-02-2020-1483/1.}}
%
\titlerunning{Finding of Multiple Effective Decisions in MOO Problems}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Konstantin Barkalov\orcidID{0000-0001-5273-2471} \and
Victor Gergel\orcidID{0000-0002-4013-2329} \and
Vladimir Grishagin \and
Evgeniy Kozinov\orcidID{0000-0001-6776-0096} 
}
%
\authorrunning{K. Barkalov, V. Gergel, V. Grishagin, E. Kozinov}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia 
\email{\{konstantin.barkalov,evgeny.kozinov\}@itmm.unn.ru,\{gergel,vagris\}@unn.ru}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
This paper considers computationally intensive multicriteria optimization problems which require computing multiple Pareto-optimal decisions. It is also assumed that efficiency criteria may be multiextremal, and the cost of calculating function values may be quite high. The proposed approach is based on the reduction of multicriteria optimization problems to one-dimensional global optimization problems that can be solved using effective statistical information algorithms of global search. One key innovation of the developed approach consists in the possibility of solving several global optimization problems simultaneously, which allows multiple Pareto-optimal decisions to be obtained. Besides, such approach provides for reuse of the computed search information, which considerably reduces computational effort for solving multicriteria optimization problems. Computational experiments confirm the potential of the proposed approach.

@@

Choosing the best options (decision making) in situations with many different alternatives is a problem that occurs in almost every domain of human activity. In the simplest cases, decision making problems can be viewed as optimization problems of various kinds such as convex programming, discrete optimization, nonlinear programming, etc. In more complex situations, the objective functions that determine the effectiveness of decisions can be multiextremal, and decision making will require solving global optimization problems. In the most general case, however, the effectiveness of decisions may be determined by several objective criteria, which necessitates solving multi-objective optimization (MOO) problems. Thus, MOO problems are the most common decision making settings that one has to solve in many scientific and technical applications. A large number of approaches and methods have been developed to solve such problems and they have been used to solve many decision making problems in various fields of practical applications -- see e.g. \cite{c1,c2,c3,c4,c5,c6,c7,c8,c9}.

@@

One of the approaches most commonly used in the search for efficient solutions is scalarization of the vector efficiency criterion, when a MOO problem is reduced to solving one or several scalar (in the general case, global) optimization problems in which the objective function is the only function. The family of optimization problems generated in such approach is further referred to as multiple global optimization problem (MGO). Within this approach, we can distinguish lexicographic optimization methods where objective functions are ordered by importance, thus  allowing the optimization of the functions to be carried out sequentially as their importance decreases [10]. Possible scalarization methods also include various methods of the efficiency criteria convolution such as the weighted sum method, the compromise programming method, the reference point method, the weighted min-max method, etc. -- see, for example, \cite{c2,c11,c12}.

@@

The structure of the paper is as follows. In Section \ref{sec:02}, we give the statement of the multicriteria optimization problem. Section \ref{sec:03} presents a scheme for reduction of multicriteria optimization problems to one-dimensional global optimization problems as well as some methods for solving such problems. Section \ref{sec:04} considers the proposed approach for simultaneous solution of several global optimization problems yielding several Pareto-optimal decisions at a time. Section \ref{sec:05} contains the results of numerical experiments confirming the effectiveness of the proposed approach. In conclusion, the results obtained are discussed and possible main lines of action for further research are outlined.

@@

Without loss of generality, it is assumed that objective functions should be minimized to improve the decision efficiency $y \in D$. It is also assumed that $f_i(y)$, $1 \leq i \leq s$, are multiextremal and have the form of time-consuming ``black-box'' computational procedures. It is also assumed that the objective functions. It is also assumed that the objective functions $f_i(y)$, $1 \leq i \leq s$, satisfy the Lipschitz condition 
\begin{equation}
\label{eq:03}
|f_i(y') - f_i(y'')| \leq L_i\|y' - y''\|, y', y'' \in D, 1 \leq i \leq s,
\end{equation}
where $L_i$, $1 \leq i \leq s$, are the Lipschitz constants and $\|*\|$ denotes the Euclidean norm in $R^N$.  Condition (\ref{eq:03}) means that for small variations of the parameter $y \in D$, the corresponding changes in the values of the functions $f_i(y)$, $1 \leq i \leq s$, are bounded.

@@

\textbf{1. Scalarization of multiple objective functions.} In the most general form, the global optimization problem generated by the scalarization of multiple objective functions of the MOO problem can be represented in the form
\begin{equation}
\label{eq:04}
\min {\varphi(y)} = F(\lambda, y), y \in D,
\end{equation}
where $F$ is a scalar multiextremal function generated as a result of scalarization of objective functions $f_i$, $1 \leq i \leq s$, $\lambda$ is the vector of parameters of the applied convolution of functions, and $D$ is the search domain from (\ref{eq:01}). In the proposed approach, we use for scalarization the compromise programming method \cite{c2,c11}, where solution of the MOO problem consists in finding the efficient decision corresponding most closely to the optimality measures of the specified reference decision $y^0 \in D$. In this case, a possible variant of convolution of objective functions $f_i$, $1 \leq i \leq s$, can have the form  
\begin{equation}
\label{eq:05}
\min F(\lambda,y)=1/s \sum_{i=1}^s {\lambda_i (f_i (y)-f_i (y^0 ))^2}, y \in D,
\end{equation}
where $F(\lambda,y)$ is the standard deviation of the values of the objective functions $f_i$, $1 \leq i \leq s$, for the decision $y \in D$ and for the specified reference decision $y^0 \in D$, while the coefficients $0 \leq \lambda_i \leq 1$, $1 \leq i < s$, are measures of the significance of the approximation accuracy for each variable $y_i$, $1 \leq i \leq N$, separately  without loss of generality, we can assume that the area of possible values of the coefficients $\lambda$ is a set 
\begin{equation}
\label{eq:06}
\lambda=(\lambda_1,\lambda_2, \dots, \lambda_s)\in \Lambda \subset R^s: \sum_{i=1}^{s}\lambda_i=1, \lambda_i \geq 0, 1 \leq i \leq s.
\end{equation}

The reference decision $y^0 \in D$ in (\ref{eq:05}) can be known \textit{a priori} or determined on the basis of some known prototype. In many cases, an abstract ideal decision $y^0 \in D$ is used as the reference decision $y^*\in D$ in which the objective functions $f_i$, $1 \leq i \leq s$, assume minimal possible values, i.e.
\begin{equation}
\label{eq:07}
f_i^* =f_i (y^*)=\min_{y \in D} f_i (y), 1 \leq i \leq s.
\end{equation}

@@

The resulting one-dimensional functions $\varphi(y(x))$ satisfy the uniform H{\" o}lder condition, i.e.
\begin{equation}
\label{eq:10}
|\varphi (y(x'))- \varphi (y(x''))| \leq H |x'-x''|^{1/N},
\end{equation}

where the constant $H$ is defined by the relation $H=2L\sqrt{N+3}$, $L$ is the Lipschitz constant from (\ref{eq:08}), and $N$ is the dimensionality of the optimization problem (\ref{eq:01}).

As a result of dimensionality reduction, the computational scheme for solving problem (\ref{eq:09}) consists of the following (see Fig. \ref{fig:1}):

\begin{itemize}
	\item The optimization algorithm minimizes the reduced one-dimensional function $\varphi(y(x))$ from (\ref{eq:09}),
	\item After determining the point $x \in [0,1]$ of the next iteration of the one-dimensional global search, the multi-dimensional image $y \in D$ is computed for the mapping $y(x)$,
	\item At the multi-dimensional point $y \in D$, the value of the original multi-dimensional function $\varphi(y)$ is computed from (\ref{eq:04}),
	\item The computed value $z = \varphi(y)$ is then used as the value of the reduced one-dimensional function $\varphi(y(x))$ at  the point $x\in[0,1]$.
\end{itemize}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{fig1}
  \caption{General scheme for solving the problem (\ref{eq:09})}
  \label{fig:1}
\end{figure}

\textbf{3. Solving the one-dimensional reduced optimization problem.} Using the dimensionality reduction results in one more additional advantage of the proposed approach: many well-known one-dimensional global search algorithms (possibly, after some additional generalisation) can be used to solve the initial multidimensional MOO problem from (\ref{eq:01}) -- see e.g. \cite{c22,c23,c24,c25,c26,c27,c28,c29}. At the same time, it should be noted that most works where dimensionality reduction is a key feature for solving multiextremal problems like (\ref{eq:04}) rely on the statistical information theory of global search \cite{c15}. This theory has provided the a basis for developing a large number of effective methods for multi-extremal optimization -- see, e.g., \cite{c28,c29,c30,c31,c32,c33,c34,c35,c36,c37}.

Within the framework of statistical information theory, a general computational scheme of global optimization algorithms was proposed which in brief is as follows \cite{c15,c30,c35}.

@@

This estimate is determined by means of characteristics of intervals $R(i)$, $1 \leq i<k$, whose values should be proportional to the degree of possibility that the global minimum is located  in these intervals. The type of these characteristics depends on the global optimization algorithm used -- thus, for example, for an algorithm which constructs a uniform dense grid in the global search area, the characteristic is simply the interval length 
\begin{equation}
\label{eq:12}
R(i)=(x_i-x_{i-1}), 1 \leq i<k.
\end{equation}

For the algorithms proposed in \cite{c22,c23}, when dimensionality reduction is applied, the characteristic is an estimate of the minimum possible value of the function $\varphi(y(x))$ to be minimized on the interval $(x_{i-1},x_i )$, $1 \leq i < k$, i.e.
\begin{equation}
\label{eq:13}
R(i) = 0.5 H \varrho_i - 0.5 (z_{i-1} + z_i ), \varrho_i = (x_i-x_{i-1} ) ^ {1⁄N}, 1 \leq i < k,
\end{equation}
where $H$ is the H{\" o}lder constant from (\ref{eq:10}) for the reduced global optimization problem being solved (\ref{eq:09}), $z_i = \varphi(y(z_i))$, $1 \leq i < k$ and $N$ is the dimensionality of the problem from (\ref{eq:01})).  For the global search algorithm (GSA) \cite{c15,c24} developed in the framework of the statistical information approach, the characteristic is
\begin{equation}
\label{eq:14}
R(i)=m\varrho_i+\frac{z_i-z_{i-1}^2}{m\varrho_i}-2(z_{i-1}+z_i ),\varrho_i=(x_i-x_{i-1})^{1/N}, 1 \leq i < k,
\end{equation}
where $m$ is a numerical estimate of the H{\" o}lder constant derived from available search information
\begin{equation}
\label{eq:15}
m= r M, M = \max |z_i-z_{i-1} | ⁄ \varrho_i, 1 \leq i<k
\end{equation}
 ($r>1$ is a parameter of the GSA algorithm).

@@

\textit{Step 2.} Select the next iteration point in the interval with the maximum characteristic
\begin{equation}
\label{eq:17}
x^{k+1}=X(x_{t-1}, x_t),
\end{equation}
calculate the value $z^{k+1}$ of the function to be minimized at this point (the procedure for calculating the function value will be further referred to as a trial).

\textit{Step 3.} Check the stopping condition
\begin{equation}
\label{eq:18}
\varrho_t \leq \varepsilon,
\end{equation}
where $\varrho_t = (x_t-x_{t-1})^{1⁄N}$, $t$ from (\ref{eq:16}) and $\varepsilon > 0$ is the specified accuracy of the solution. If the stopping condition (\ref{eq:18}) is reached, then the solving of the optimization problem is stopped, otherwise $k=k+1$ is assumed and the next global search iteration begins.

@@

It should be noted again that the computational scheme discussed above is quite general. Many global search algorithms can be represented within this scheme, as evidenced, in particular, by examples (\ref{eq:12})-(\ref{eq:14}) and other characteristic algorithms -- see, for example, \cite{c28,c29,c30,c31,c32,c33,c34,c35,c36,c37}.

Convergence conditions for characteristic algorithms depend on the properties of interval characteristics used. One of sufficient conditions for convergence of algorithms is, for example, the requirement that the characteristic of the interval containing the global minimum point should take on the maximum value at step 1 of the characteristic scheme during the global search iterations. This condition is satisfied, for example, for the multidimensional generalized algorithms proposed in \cite{c22,c23} when the H{\"o}lder constant from (\ref{eq:10}) is specified exactly. For GSA, a sufficient condition for convergence is the relation \cite{c15}
\begin{equation}
\label{eq:20}
m \geq 2^{3-1⁄N} L\sqrt{N+3},
\end{equation}
which must be fulfilled starting from some iteration $k>1$ of the global search ($L$ is the Lipschitz constant from (\ref{eq:08})). Moreover, if condition (\ref{eq:20}) is satisfied, only the points of the global minimum of the function $\varphi(y)$ from (\ref{eq:04}) will be the limit points of the trial sequence $\{y^k=y(x^k)\}$ generated by the GSA algorithm.

\section{An Approach for Simultaneous Finding of Multiple Effective Solutions in Multi-objective Optimization Problems} \label{sec:04}

\textbf{1. The need to solve multiple global optimization problems.} As mentioned earlier, in the process of solving the MOO problem, it may be necessary to find several different effective decisions due to possible changes in the optimality requirements. Obtaining different  effective decisions in the proposed approach is ensured by choosing different convolution coefficients (importance indicators) for the objective functions $f_i(y)$, $1 \leq i \leq s$, which results in obtaining different  scalar multi-extremal functions $F(\lambda, y)$ from (\ref{eq:04}). The resulting different functions $F(\lambda, y)$ for different values of the coefficients $\lambda \in \Lambda$ can be optimized sequentially. This deermines a multi-stage scheme for solving the MOO problem, when at each stage DM specifies the necessary scalarization coefficients, then the resulting scalar optimization problem is solved, after which DM analyzes the effective solution found.

The general scheme considered above can be extended by the option of selecting not one but several different scalarization coefficients at each stage. Having such an option makes it easier for DM  to specify importance coefficients $\lambda \in \Lambda$ of the objective functions $f_i(y)$, $1 \leq i \leq s$. By solving simultaneously several generated scalar optimization problems, one can  obtain the estimates of efficient decisions at the earliest stages of computing, which makes it possible to dynamically (in the process of computation) change the set of problems being solved -- to stop solving those that obviously have no prospect of success (from DM's point of view) or to add new optimization problems.

Such generalization of the process of solving a MOO problem means that at each current moment of calculations there is a set of functions being optimized simultaneously having the form (\ref{eq:04})
\begin{equation}
\label{eq:21}
\Phi_p (y)=\{ F(\lambda_1,y), F(\lambda_2,y),\dots,F(\lambda_p,y)  \},  \lambda_i \in \Delta, 1 \leq i \leq p
\end{equation}
which can be changed dynamically in the course of the calculations by adding new or removing existing optimization functions $F(\lambda, y)$ from (\ref{eq:04}).

\textbf{2. Step-by-step solution of a set of global optimization problems.} As shown in Section \ref{sec:03}, the statistical information multiextremal optimization algorithms used in the proposed approach determine the points of consecutive iterations of the global search taking into account the search information
\begin{equation}
\label{eq:22}
A_k=\{(x_i,z_i,f_i )^T:1 \leq i \leq k\},
\end{equation}
obtained in the calculations (see (\ref{eq:14})-(\ref{eq:17})). In (\ref{eq:22}) $x_i$, $1 \leq i \leq k$,  are the reduced points of performed global search iterations ordered in ascending order of coordinates,  $z_i$, $f_i$, $1 \leq i \leq k$, are values of the scalar function $F(\lambda, y)$, from (\ref{eq:04}) and objective functions $f_i(y)$, $1 \leq i \leq s$, from (\ref{eq:01}) of the current optimization problem to be solved at the points $x_i=y(x_i)$, $1 \leq i \leq k$. By using search information $A_k$ from (\ref{eq:22}) when choosing next search iterations it is possible to solve global optimization problems more efficiently and to provide convergence of algorithms only to the global minima of multiextremal functions being minimized.

It is important to note that since the set $\Phi_t(y)$ of functions being optimized  simultaneously is generated from the same MOO problem from (\ref{eq:01}), the existence of the set $A_k$ from (\ref{eq:12}) allows us to adjust the results of all previously performed calculations of the values of the objective functions $f_i(y)$, $1 \leq i \leq s$, to the values of the next optimized function $F(\lambda, y)$ from (\ref{eq:04}) without repeating any time consuming calculations of values, i.e.
\begin{equation}
\label{eq:23}
(x_i,f_i ) \to z_i=F(\lambda, y(x_i)) ,1 \leq i \leq k.
\end{equation}

Thus, all the search information $A_k$ from (\ref{eq:22}), recalculated according to (\ref{eq:23}), can be reused to continue solving the problems of the set $\Phi_t(y)$. Such a possibility provides a significant reduction in computations up to performing only a limited set of global search iterations (see the results of computational experiments in Section \ref{sec:03}).

     This type of information connectivity of functions in the set $\Phi_t(y)$ from (\ref{eq:21}) makes it possible to generalize the computational scheme (\ref{eq:16})-(\ref{eq:18}) for solving a single global optimization problem for the case of optimizing the functions of the set $\Phi_t(y)$ from (\ref{eq:21}) by adding a preliminary step of the search information  transformation.

\textit{Step 0.} Adjust the state of search information $A_k$ from (\ref{eq:22}) to the values of the function $F(\lambda, y)$ from the set $\Phi_t(y)$ according to rule (\ref{eq:23}).

The GSA algorithm applied to optimize the functions of the set $\Phi_t(y)$ from (\ref{eq:21}) and using search information $A_k$ will be further referred to as the Multiple Global Search Algorithm (MGSA).

>>>>

\begin{document}
%
\title{An Approach for Simultaneous Finding of Multiple Effective Decisions in Multi-objective Optimization Problems
%with non-convex constraints 
\thanks{This work was supported by the Ministry of Science and Higher Education of the Russian Federation, project no.0729-2020-0055, and by the Research and Education Mathematical Center, project no. 075-02-2020-1483/1.}}
%
\titlerunning{Finding of Multiple Effective Decisions in MOO Problems}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Konstantin Barkalov\orcidID{0000-0001-5273-2471} \and
Victor Gergel\orcidID{0000-0002-4013-2329} \and
Vladimir Grishagin \and
Evgeniy Kozinov\orcidID{0000-0001-6776-0096} 
}
%
\authorrunning{K. Barkalov, V. Gergel, V. Grishagin, E. Kozinov}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia 
\email{\{konstantin.barkalov,evgeny.kozinov\}@itmm.unn.ru,\{gergel,vagris\}@unn.ru}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
This paper considers computationally intensive multi-objective optimization problems which require computing multiple Pareto-optimal decisions. It is also assumed that efficiency criteria may be multiextremal, and the cost of calculating function values may be quite high. The proposed approach is based on the reduction of multi-objective optimization problems to one-dimensional global optimization problems that can be solved using effective information-statistical algorithms of global search. One key innovation of the developed approach consists in the possibility of solving several global optimization problems simultaneously, which allows multiple Pareto-optimal decisions to be obtained. Besides, such approach provides for reuse of the computed search information, which considerably reduces computational effort for solving multi-objective optimization problems. Computational experiments confirm the potential of the proposed approach.

@@

Choosing the optimal options (decision making) in situations with many different alternatives is a problem that occurs in almost every domain of human activity. In the simplest cases, decision making problems can be viewed as optimization problems of various kinds such as convex programming, discrete optimization, nonlinear programming, etc. In more complex situations, the objective functions that determine the effectiveness of decisions can be multiextremal, and decision making will require solving global optimization problems. In the most general case, however, the effectiveness of decisions may be determined by several objective criteria, which necessitates solving multi-objective optimization (MOO) problems. Thus, MOO problems are the most common decision making statements that one has to solve in many scientific and technical applications. A large number of approaches and methods have been developed to solve such problems and they have been used to solve many decision making problems in various fields of practical applications -- see e.g. \cite{c1,c2,c3,c4,c5,c6,c7,c8,c9}.

@@

One of the approaches most commonly used in the search for efficient decisions is scalarization of the vector efficiency criterion, when a MOO problem is reduced to solving one or several scalar (in the general case, global) optimization problems in which the objective functions are converted to the integrated scalar function. The family of optimization problems generated in such approach is further referred to as multiple global optimization problem (MGO). Within this approach, we can distinguish lexicographic optimization methods where objective functions are ordered by importance, thus  allowing the optimization of the functions to be carried out sequentially as their importance decreases [10]. Possible scalarization methods also include various methods of the efficiency criteria convolution such as the weighted sum method, the compromise programming method, the reference point method, the weighted min-max method, etc. -- see, for example, \cite{c2,c11,c12}.

@@

The structure of the paper is as follows. In Section \ref{sec:02}, we give the statement of the multi-objective optimization problem. Section \ref{sec:03} presents a scheme for reduction of multi-objective optimization problems to one-dimensional global optimization problems as well as some methods for solving such problems. Section \ref{sec:04} considers the proposed approach for simultaneous solution of several global optimization problems yielding several Pareto-optimal decisions at a time. Section \ref{sec:05} contains the results of numerical experiments confirming the effectiveness of the proposed approach. In conclusion, the results obtained are discussed and possible main lines of action for further research are outlined.

@@

Without loss of generality, it is assumed that objective functions should be minimized to improve the decision efficiency $y \in D$. It is also assumed that $f_i(y)$, $1 \leq i \leq s$, are multiextremal and have the form of time-consuming ``black-box'' computational procedures. It is also assumed that the objective functions $f_i(y)$, $1 \leq i \leq s$, satisfy the Lipschitz condition 
\begin{equation}
\label{eq:03}
|f_i(y') - f_i(y'')| \leq L_i\|y' - y''\|, y', y'' \in D, 1 \leq i \leq s,
\end{equation}
where $L_i$, $1 \leq i \leq s$, are the Lipschitz constants and $\|*\|$ denotes the Euclidean norm in $R^N$.  Condition (\ref{eq:03}) means that for small variations of the parameter $y \in D$, the corresponding changes in the values of the functions $f_i(y)$, $1 \leq i \leq s$, are bounded.

@@

\textbf{1. Scalarization of multiple objective functions.} In the most general form, the global optimization problem generated by the scalarization of multiple objective functions of the MOO problem can be represented in the form
\begin{equation}
\label{eq:04}
\min {\varphi(y)} = F(\lambda, y), y \in D,
\end{equation}
where $F$ is a scalar multiextremal function generated as a result of scalarization of objective functions $f_i$, $1 \leq i \leq s$, $\lambda$ is the vector of parameters of the applied convolution of functions, and $D$ is the search domain from (\ref{eq:01}). In the proposed approach, we use for scalarization the compromise programming method \cite{c2,c11}, where solution of the MOO problem consists in finding the efficient decision corresponding most closely to the values of the objective functions of the specified reference decision $y^0 \in D$. In this case, a possible statement of the statement scalar optimization problem can have the form 
\begin{equation}
\label{eq:05}
\min F(\lambda,y)=\min 1/s \sum_{i=1}^s {\lambda_i (f_i (y)-f_i (y^0 ))^2}, y \in D,
\end{equation}
where $F(\lambda,y)$ is the standard deviation of the values of the objective functions $f_i$, $1 \leq i \leq s$, for the decision $y \in D$ and for the specified reference decision $y^0 \in D$, while the coefficients $0 \leq \lambda_i \leq 1$, $1 \leq i < s$, are measures of the significance of the approximation accuracy for each variable $y_i$, $1 \leq i \leq N$, separately. Without loss of generality, we can assume that the area of possible values of the coefficients $\lambda$ is a set 
\begin{equation}
\label{eq:06}
\lambda=(\lambda_1,\lambda_2, \dots, \lambda_s)\in \Lambda \subset R^s: \sum_{i=1}^{s}\lambda_i=1, \lambda_i \geq 0, 1 \leq i \leq s.
\end{equation}

The reference decision $y^0 \in D$ in (\ref{eq:05}) can be known \textit{a priori} or determined on the basis of some known prototype. In many cases, an abstract ideal decision $y^0 \in D$ is used as the reference decision $y^*\in D$ in which the objective functions $f_i$, $1 \leq i \leq s$, have minimal possible values, i.e.
\begin{equation}
\label{eq:07}
f_i^* =f_i (y^*)=\min_{y \in D} f_i (y), 1 \leq i \leq s.
\end{equation}

@@

The resulting one-dimensional functions $\varphi(y(x))$ satisfy the uniform H{\" o}lder condition, i.e.
\begin{equation}
\label{eq:10}
|\varphi (y(x'))- \varphi (y(x''))| \leq H |x'-x''|^{1/N},
\end{equation}
where the constant $H$ is defined by the relation $H=2L\sqrt{N+3}$, $L$ is the Lipschitz constant from (\ref{eq:08}), and $N$ is the dimensionality of the optimization problem (\ref{eq:01}).

\textbf{3. Solving the one-dimensional reduced optimization problem.} Using the dimensionality reduction results in one more additional advantage of the proposed approach: many well-known one-dimensional global search algorithms (possibly, after some additional generalisation) can be used to solve the initial multidimensional MOO problem from (\ref{eq:01}) -- see e.g. \cite{c22,c23,c24,c25,c26,c27,c28,c29}. At the same time, it should be noted that most works where dimensionality reduction is a key feature for solving multiextremal problems like (\ref{eq:04}) rely on the information-statistical theory of global search \cite{c15}. This theory has provided the a basis for developing a large number of efficient methods for multi-extremal optimization -- see, e.g., \cite{c28,c29,c30,c31,c32,c33,c34,c35,c36,c37}.

Within the framework of information-statistical theory, a general computational scheme of global optimization algorithms was proposed which in brief is as follows \cite{c15,c30,c35}.

@@

This estimate is determined by means of characteristics of intervals $R(i)$, $1 \leq i<k$, whose values should be proportional to the degree of possibility that the global minimum is located  in these intervals. The type of these characteristics depends on the global optimization algorithm used -- thus, for example, for an algorithm which constructs a uniform dense grid in the global search domain, the characteristic is simply the interval length 
\begin{equation}
\label{eq:12}
R(i)=(x_i-x_{i-1}), 1 \leq i<k.
\end{equation}

For the algorithms proposed in \cite{c22,c23}, when dimensionality reduction is applied, the characteristic is an estimate of the minimum possible value of the function $\varphi(y(x))$ to be minimized on the interval $(x_{i-1},x_i )$, $1 \leq i < k$, i.e.
\begin{equation}
\label{eq:13}
R(i) = 0.5 H \varrho_i - 0.5 (z_{i-1} + z_i ), \varrho_i = (x_i-x_{i-1} ) ^ {1⁄N}, 1 \leq i < k,
\end{equation}
where $H$ is the H{\" o}lder constant from (\ref{eq:10}) for the reduced global optimization problem being solved (\ref{eq:09}), $z_i = \varphi(y(z_i))$, $1 \leq i < k$ and $N$ is the dimensionality of the problem from (\ref{eq:01})).  For the global search algorithm (GSA) \cite{c15,c24} developed in the framework of the information-statistical approach, the characteristic is
\begin{equation}
\label{eq:14}
R(i)=m\varrho_i+\frac{z_i-z_{i-1}^2}{m\varrho_i}-2(z_{i-1}+z_i ),\varrho_i=(x_i-x_{i-1})^{1/N}, 1 \leq i < k,
\end{equation}
where $m$ is a numerical estimate of the H{\" o}lder constant derived from available search information
\begin{equation}
\label{eq:15}
m= r M, M = \max |z_i-z_{i-1} | ⁄ \varrho_i, 1 \leq i<k
\end{equation}
 ($r>1$ is a parameter of the GSA algorithm).

@@

\textit{Step 2.} Select the next iteration point in the interval with the maximum characteristic
\begin{equation}
\label{eq:17}
x^{k+1}=X(x_{t-1}, x_t)\footnote{The rule $X$ for selecting the point $x^{k+1}$ of the next iteration in the interval $(x_{t-1}, x_t)$ is stated by the global optimization algorithm},
\end{equation}
calculate the value $z^{k+1}$ of the function to be minimized at this point (the procedure for calculating the function value will be further referred to as a trial).

\textit{Step 3.} Check the stopping condition
\begin{equation}
\label{eq:18}
\varrho_t \leq \varepsilon,
\end{equation}
where $\varrho_t = (x_t-x_{t-1})^{1⁄N}$, $t$ from (\ref{eq:16}) and $\varepsilon > 0$ is the specified accuracy of the solution. If the stopping condition (\ref{eq:18}) is met, then the solving of the optimization problem is stopped, otherwise $k=k+1$ is assumed and the next global search iteration begins.

@@

It should be noted again that the computational scheme discussed above is quite general. Many global search algorithms can be represented within this scheme, as evidenced, in particular, by examples (\ref{eq:12})-(\ref{eq:14}) and other global optimization algorithms -- see, for example, \cite{c28,c29,c30,c31,c32,c33,c34,c35,c36,c37}.

Convergence conditions of the algorithms formulated in the form (\ref{eq:16})-(\ref{eq:18}). One of sufficient conditions for convergence of algorithms is, for example, the requirement that the characteristic of the interval containing the global minimum point should take on the maximum value at step 1 of the scheme (\ref{eq:16})-(\ref{eq:18}) during the global search iterations. This condition is satisfied, for example, for the multidimensional generalized algorithms proposed in \cite{c22,c23} when the H{\"o}lder constant from (\ref{eq:10}) is specified exactly. For GSA, a sufficient condition for convergence is the relation \cite{c15}
\begin{equation}
\label{eq:20}
m \geq 2^{3-1⁄N} L\sqrt{N+3},
\end{equation}
which must be fulfilled starting from some iteration $k>1$ of the global search ($L$ is the Lipschitz constant from (\ref{eq:08})). Moreover, if condition (\ref{eq:20}) is satisfied, only the points of the global minimum of the function $\varphi(y)$ from (\ref{eq:04}) will be the limit points of the trial sequence $\{y^k=y(x^k)\}$ generated by the GSA algorithm.

\section{An Approach for Simultaneous Finding of Multiple Efficient Decisions in Multi-objective Optimization Problems} \label{sec:04}

\textbf{1. The need to solve multiple global optimization problems.} As mentioned earlier, in the process of solving the MOO problem, it may be necessary to find several different efficient decisions due to possible changes in the optimality requirements. Obtaining different efficient decisions in the proposed approach is ensured by choosing different convolution coefficients (importance indicators) for the objective functions $f_i(y)$, $1 \leq i \leq s$, which results in obtaining different  scalar multi-extremal functions $F(\lambda, y)$ from (\ref{eq:04}). These functions $F(\lambda, y)$ for different values of the coefficients $\lambda \in \Lambda$ can be optimized sequentially. This determines a multi-stage scheme for solving the MOO problem, when at each stage DM specifies the necessary scalarization coefficients, then the resulting scalar optimization problem is solved, after which DM analyzes the efficient decisions found.

The general scheme considered above can be extended by the option of selecting not one but several different scalarization coefficients at each stage. Having such an option makes it easier for DM  to specify importance coefficients $\lambda \in \Lambda$ of the objective functions $f_i(y)$, $1 \leq i \leq s$. By solving simultaneously several generated scalar optimization problems, one can  obtain the estimates of efficient decisions at the earliest stages of computing, which makes it possible to dynamically (in the process of computations) change the set of problems being solved -- to stop solving those that obviously have no prospect of success (from DM's point of view) or to add new optimization problems.

Such generalization of the process of solving a MOO problem means that at each current moment of calculations there is a set of functions being optimized simultaneously having the form (\ref{eq:04})
\begin{equation}
\label{eq:21}
\Phi_p (y)=\{ F(\lambda_1,y), F(\lambda_2,y),\dots,F(\lambda_p,y)  \},  \lambda_i \in \Lambda, 1 \leq i \leq p
\end{equation}
which can be changed dynamically in the course of the calculations by adding new or removing existing optimization functions $F(\lambda, y)$ from (\ref{eq:04}).

\textbf{2. Step-by-step solution of a set of global optimization problems.} As shown in Section \ref{sec:03}, the information-statistical multiextremal optimization algorithms used in the proposed approach determine the points of consecutive iterations of the global search taking into account the search information
\begin{equation}
\label{eq:22}
A_k=\{(x_i,z_i,f_i )^T:1 \leq i \leq k\},
\end{equation}
obtained in the calculations (see (\ref{eq:14})-(\ref{eq:17})). In (\ref{eq:22}) $x_i$, $1 \leq i \leq k$,  are the reduced points of performed global search iterations ordered in ascending order of coordinates,  $z_i$, $f_i$, $1 \leq i \leq k$, are values of the scalar function $F(\lambda, y)$, from (\ref{eq:04}) and objective functions $f_i(y)$, $1 \leq i \leq s$, from (\ref{eq:01}) of the current optimization problem to be solved at the points $x_i=y(x_i)$, $1 \leq i \leq k$. By using search information $A_k$ from (\ref{eq:22}) when choosing next search iterations it is possible to solve global optimization problems more efficiently and to provide convergence of algorithms only to the global minima of multiextremal functions being minimized.

It is important to note that since the set $\Phi_p(y)$ of functions being optimized  simultaneously is generated from the same MOO problem from (\ref{eq:01}), the existence of the set $A_k$ from (\ref{eq:12}) allows us to adjust the results of all previously performed calculations of the values of the objective functions $f_i(y)$, $1 \leq i \leq s$, to the values of the next optimized function $F(\lambda, y)$ from (\ref{eq:04}) without repeating any time consuming calculations of values, i.e.
\begin{equation}
\label{eq:23}
(x_i,f_i ) \to z_i=F(\lambda, y(x_i)) ,1 \leq i \leq k.
\end{equation}

Thus, all the search information $A_k$ from (\ref{eq:22}), recalculated according to (\ref{eq:23}), can be reused to continue solving the problems of the set $\Phi_p(y)$. Such a possibility provides a significant reduction in computations up to performing only a limited set of global search iterations (see the results of numerical experiments in Section \ref{sec:03}).

     This type of information connectivity of functions in the set $\Phi_p(y)$ from (\ref{eq:21}) makes it possible to generalize the computational scheme (\ref{eq:16})-(\ref{eq:18}) for solving a single global optimization problem for the case of optimizing the functions of the set $\Phi_p(y)$ from (\ref{eq:21}) by adding a preliminary step of the search information  transformation.

\textit{Step 0.} Adjust the state of search information $A_k$ from (\ref{eq:22}) to the values of the function $F(\lambda, y)$ from the set $\Phi_p(y)$ according to rule (\ref{eq:23}).

The GSA algorithm applied to optimize the functions of the set $\Phi_p(y)$ from (\ref{eq:21}) and using the search information $A_k$ will be further referred to as the Multiple Global Search Algorithm (MGSA).
